SMT-Boosted Security Types for Low-Level MPC

### 2024-10-14 esop25/cpj.tex
checkpoint

For example, given some string $y$ and a message send such as:
$$
\xassign{\mx{w}{1}}{\secret{w} \ftimes \flip{w}}{2}
$$
we would assign the following type to denote that $\mx{w}{1}$ is dependent on
both $\sx{w}{2}$ and $\rx{w}{2}$:
$$
\mx{w}{1} : \setit{\sx{w}{2}, \rx{w}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{w}{1}}{\secret{w} \fminus \flip{w}}{2}
$$
we want the type of $\mx{w}{1}$ to be an ``encrypted'' type
that is by itself independent of $\sx{w}{2}$. However, we also
want to track the dependence of $\mx{w}{2}$ and $\rx{w}{2}$.
In short, we assign $\mx{w}{1}$ a ``ciphertext type'', aka
a \emph{ciphertype}:
$$
\mx{w}{1} : \cty{\rx{w}{2}}{\setit{\sx{w}{2}}}
$$
indicating syntactically that $\mx{w}{1}$ is an encryption of a
value dependent on $\sx{w}{2}$ with $\rx{w}{2}$ and as long as client
1 does not subsequently receive messages that are dependent on
$\rx{w}{2}$, the value $\sx{w}{2}$ remains hidden.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and shares
its secret input encoding with the ``evaluator'' by using
it as a selection bit. That is, assuming that client 2 is the garbler and
client 1 is the evaluator, we can define:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertype to $\mx{w}{1}$ as
above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

@@

The syntax of types $\ty$ and type environments $\Gamma$ are presented
in Figure \ref{fig-cpj}\footnote{In general we use $2^{\chi}$ to denote
the powerset of terms in any grammatical sort $\chi$.}. We also
define type judgements for expressions and programs by direct
interpretation as terms $\phi$ and constraints $\eqs$ respectively,
via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule captures
the use ``fresh'' samples for encryption discussed above. Derivations
enforce linearity by requiring that variables used for encryption
are added to judgements, and guaranteed to not be used elsewhere
via a form of union that ensures disjointness of combined sets:
\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ if $R_1 \cap R_2 = \varnothing$ and is otherwise
  undefined.
\end{definition}
The \TirName{DepTy} rule handles all other syntactic forms and just
records the dependencies on variables in the term.  The \TirName{Send}
rule records the types of messages, reveals, and outputs in
enviroments, and the \TirName{Seq} rule combines the types of
sequenced programs.  Given this we define validity of program type
judgements as follows. 
\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}
It is important to note that this presentation of derivations is
logical, not algorithmic-- in particular it is not syntax-directed
due to the \TirName{Encode} rule.

@@

    \eqj{R_2}{\eqs}{R_2}{\ty_1}
  \end{mathpar}
  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
  $\vars{\phi}$ in $M'$ 
  \cite{barthe2019probabilistic}. Observing that:
  $$
  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
  $$
  The result in this subcase follows by the (second) induction hypothesis. \qed
\end{proof}

>>>>

For example, given some string $y$ and a message send such as:
$$
\xassign{\mx{w}{1}}{\secret{w} \ftimes \flip{w}}{2}
$$
we would assign the following type to denote that $\mx{w}{1}$ is dependent on
both $\sx{w}{2}$ and $\rx{w}{2}$:
$$
\mx{w}{1} : \setit{\sx{w}{2}, \rx{w}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{w}{1}}{\secret{w} \fminus \flip{w}}{2}
$$
we want the type of $\mx{w}{1}$ to be an ``encrypted'' type that is
by itself independent of $\sx{w}{2}$. However, we also want to track
the \emph{joint} dependence of $\mx{w}{2}$ on $\rx{w}{2}$ and
$\sx{w}{2}$, in case information about $\rx{w}{2}$ is subsequently
leaked.  In short, we assign $\mx{w}{1}$ a ``ciphertext type'', aka a
\emph{ciphertype}:
$$
\mx{w}{1} : \cty{\rx{w}{2}}{\setit{\sx{w}{2}}}
$$
indicating syntactically that $\mx{w}{1}$ is an encryption of a
value dependent on $\sx{w}{2}$ with $\rx{w}{2}$.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and shares
its secret input encoding with the ``evaluator'' by using
it as a selection bit. That is, assuming that client 2 is the garbler and
client 1 is the evaluator, we can define:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$ (with $\neg$ denoting negation in $\mathbb{F}_2$):
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$.  Since this is a
one-time-pad scheme, it is also important that we ensure that samples
are used at most once for encryption. As in previous work 
\cite{darais2019language} we use type linearity to enforce this.

@@

The syntax of types $\ty$ and type environments $\Gamma$ are presented
in Figure \ref{fig-cpj}\footnote{In general we use $2^{\chi}$ to
denote the powerset of terms in any grammatical sort $\chi$.}. We also
define type judgements for expressions and programs by direct
interpretation as terms $\phi$ and constraints $\eqs$ respectively,
via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule enforces
linearity by requiring that variables used for encryption are added to
judgements, and guaranteed to not be used elsewhere via a form of
union that ensures disjointness of combined sets:
\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ if $R_1 \cap R_2 = \varnothing$ and is otherwise
  undefined.
\end{definition}
The \TirName{DepTy} rule handles all cases and just records the
dependencies on variables in the term, a conservative approximation of
dependencies.  The \TirName{Send} rule records the types of messages,
reveals, and outputs in enviroments, and the \TirName{Seq} rule
combines the types of sequenced programs.  Given this we define
validity of program type judgements as follows.
\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}
It is important to note that this presentation of derivations is
logical, not algorithmic-- in particular it is not syntax-directed
due to the \TirName{Encode} rule.

@@

    \eqj{R_2}{\eqs}{R_2}{\ty_1}
  \end{mathpar}
  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
  $\vars(\phi)$ in $M'$ 
  \cite{barthe2019probabilistic}. Observing that:
  $$
  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
  $$
  The result in this subcase follows by the (second) induction hypothesis. \qed
\end{proof}
